---
title: "Donkey"
author: "Team"
date: "5/14/2020"
output: html_document
---

## Introduction

Veterinarians give medications to diagnose, treat, or prevent illness. depending 
on the specific ailment, drugs come in different forms and are administered take 
them in many different ways. With the dangers of farmers attempting to prescribe 
dosages themselves, they depend on healthcare providers to understand the right 
way to administer them and reduce the risks. Depending on the size and 
temperament of an animal, say a dog versus a donkey, it may be easier
or harder to weigh animals directly. Over the years methods have been developed to
weigh a donkey indirectly using simple measuring devices and use these
measurements to approximate the weight of the donkey.

Estimating donkey weights data is an incredibly important task in the lives of
these farmers who depend on donkeys to deliver items ranging from building 
materials to textiles. A  donkey weight is not relatively complicated and but 
it conveys information about all aspects of drug administration, so dosages 
must be only evaluated by a trained technician. For the same reasons, there is 
interest in generating algorithms or models to characterize weights to 
potentially aid in this process.

Each measurement of a donkey  consists of the following: the BCS, Age,Sex,Length,Girth,Height,Weight, and WeightAlt. The measuremnts these 
individual components are clinically important, as these characteristics can 
help a veterinarian find abnormalities in the health of a donkey, allowing for 
crucial diagnoses. Previous studies within statistics and machine learning that
seek to classify donkey weight for varying sizes have applied least-squares
regression methods (Milner)

In our project, we borrow a linear model proposed by Milner and apply it to 
interpret the error for the estimates in the model parameters for an appropriate 
weight. To do so, we accomplished the following:




# Load in data

```{r}
load('donkeys.rda')
head(donkeys)

train = data.frame(read.table("train.csv", sep=",", header=TRUE))
test = data.frame(read.table("test.csv", sep=",", header=TRUE))

train$BCS <- as.factor(train$BCS)
train$Age <- as.factor(train$Age)
train$Sex <- as.factor(train$Sex)

test$BCS <- as.factor(test$BCS)
test$Age <- as.factor(test$Age)
test$Sex <- as.factor(test$Sex)

```

```{r}
# drop BCS = 4.5 and BCS = 1
donkeys <- subset(donkeys, BCS!=4.5 & BCS!=1)

# drop baby donkey 
donkeys <- subset(donkeys, Weight != 27)
```

```{r}
# mean weights for different donkey sexes
mean(subset(donkeys, Sex == 'stallion')$Weight)
mean(subset(donkeys, Sex == 'female')$Weight)
mean(subset(donkeys, Sex == 'gelding')$Weight)
```




# EDA

```{r}
# pairs(Weight~BCS + Age + Sex + Length + Girth + Height, data = donkeys)
pairs(Weight~Length + Girth + Height, data = train)
```


```{r}
# Distributions of Quantitative variables
hist(donkeys$Girth)
hist(donkeys$Length)
hist(donkeys$Height)
hist(donkeys$Weight)
```


# Naive full model

```{r}
m1 = lm(Weight~ BCS + Age + Sex + Length + Girth + Height, data=train)
summary(m1)
```

```{r}
# normal pairs
pairs(Weight ~ Length + Girth + Height, data = donkeys)
```


```{r}
# pairs logged 
pairs(Weight ~ log(Length) + log(Girth) + log(Height), data = donkeys)
predictions = predict.lm(m1, test, interval="confidence")
```

```{r}
residuals = predictions[,1] - test$Weight
MSE = mean(residuals^2)
MSE
```

```{r}
test <- test %>% mutate(Residuals = residuals, Predictions=predictions[,1])
ggplot(data=test, aes(x=Predictions, y=Residuals)) + geom_point() +
  geom_hline(yintercept=0,color="red") + 
  labs(title="Residuals vs. Predictions") + 
  theme(plot.title=element_text(hjust=0.5))
```

```{r}
library(olsrr)
stepwise = ols_step_both_aic(m1)
stepwise

forward = ols_step_forward_aic(m1)
forward

backward = ols_step_backward_aic(m1)
backward
```


# Getting rid of sex

```{r}
m2 = lm(Weight~ BCS + Age  + Length + Girth + Height, data=train)
summary(m2)
```


```{r}
predictions = predict.lm(m2, test, interval="confidence")
```

```{r}
residuals = predictions[,1] - test$Weight
MSE = mean(residuals^2)
MSE
```

```{r}
test <- test %>% mutate(Residuals = residuals, Predictions=predictions[,1])
ggplot(data=test, aes(x=Predictions, y=Residuals)) + geom_point() +
  geom_hline(yintercept=0,color="red") + 
  labs(title="Residuals vs. Predictions") + 
  theme(plot.title=element_text(hjust=0.5))
```

# 


```{r}
m3 = lm(Weight~ BCS + Age + log(Length) + log(Girth) + log(Height), data=train)
summary(m3)
```


```{r}
predictions = predict.lm(m3, test, interval="confidence")
```

```{r}
residuals = predictions[,1] - test$Weight
MSE = mean(residuals^2)
MSE
```

```{r}
test <- test %>% mutate(Residuals = residuals, Predictions=predictions[,1])
ggplot(data=test, aes(x=Predictions, y=Residuals)) + geom_point() +
  geom_hline(yintercept=0,color="red") + 
  labs(title="Residuals vs. Predictions") + 
  theme(plot.title=element_text(hjust=0.5))
```

# Reproducing Modeling
```{r}
train_reproducing = read.csv("train_reproducing.csv")
train_reproducing
```


```{r}
model <- lm(2*(sqrt(Weight)-1) ~ log(Length) + log(Girth), data=train_reproducing)
summary(model)
```



